\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-06-19 doc/paper.Rmd
renamed background to methods, pulled in testing setup
reading through abstract

---
title: "Adaption of the Chumbley Score to matching of bullet striation marks"
authors:
- affiliation: Department of Statistics, Iowa State University 
  name: Ganesh Krishnan
  thanks: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
- affiliation: Department of Statistics and CSAFE, Iowa State University 
  name: Heike Hofmann
biblio-style: apsr
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
blinded: 0
keywords:
- forensic science
- toolmark
- cross-correlation
- Mann-Whitney U statistic
- land engraved areas (LEAs)
- algorithm
bibliography: bibliography
abstract: One of the major problems in forensic toolmark examination is the same-source problem, i.e. the identification of whether the marks on two different objects were made by the same tool. Profilometers and confocal microscopy and other technological advances allow us to measure 3D surfaces in previously unforeseen resolution and also enable digitized images. The literature provides various methods to measure similarity of toolmarks. In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler}, which was introduced for the assessment of marks made by screwdrivers, to assess striation marks on bullet lands for same-source identification. We provide methods to identify parameters that minimize the error rates for matching of LEAs at the example of the Hamby data, sets 44 and 252 measured by NIST and CSAFE. We provide a remedial algorithm to alleviate the problem of failed tests in the method proposed by \citet{hadler}. This increases the power of the test and reduces error rates in the process. Type II error rates for the Hamby data using the adjusted method showed an improvement of more than 20\% points over the performance of the algorithm proposed by \citet{hadler} (for a Type I error of 0.05). Overall, type II error rates of the adjusted method are on average at 0.22, putting the method on similar footing as other single feature matching approaches in the literature.
---

@@

# Background

@@

# Testing setup

## The Data

@@

## Setup
Both algorithms (CS1) and (CS2) are implemented in R \citep{R}. (CS1) is available from package `toolmaRk` \citep{toolmark}, (CS2) is available from a modified version of the `toolmaRk` package available from GitHub (\url{https://github.com/heike/toolmaRk}). We applied both methods to all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.

>>>>

---
title: "Adaption of the Chumbley Score to matching of bullet striation marks"
authors:
- affiliation: Department of Statistics, Iowa State University 
  name: Ganesh Krishnan
  thanks: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
- affiliation: Department of Statistics and CSAFE, Iowa State University 
  name: Heike Hofmann
biblio-style: apsr
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
blinded: 0
keywords:
- forensic science
- toolmark
- cross-correlation
- Mann-Whitney U statistic
- land engraved areas (LEAs)
- algorithm
- signatures 
- same-source problem
bibliography: bibliography
abstract: The same-source problem remains a major challenge in forensic toolmark and firearm examination. Technological advances in surface imaging allow measurements of 3D surfaces at previously unforeseen resolutions and enable digitized imaging. Here, we investigate the applicability of the Chumbley scoring method \citep{hadler}, developed for screwdriver markings, for same-source identification of bullet striae. We provide methods to identify parameters that minimize the error rates for matching of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE. We suggest a remedial algorithm to alleviate the problem of failed tests in the method, increasing both the power of the test and reducing error rates. Type II error rates of the proposed method improve by more than one third (Type I error of 0.05) and are on average about 0.22. This puts the proposed method on similar footing as other single feature matching approaches in the literature. 
---

@@

# Methods

@@

## Testing setup

### The Data

@@

### Setup
Both algorithms (CS1) and (CS2) are implemented in R \citep{R}. (CS1) is available from package `toolmaRk` \citep{toolmark}, (CS2) is available from a modified version of the `toolmaRk` package available from GitHub (\url{https://github.com/heike/toolmaRk}). We applied both methods to all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.

### 2018-06-19 doc/paper.tex
reading through abstract

\bigskip
\begin{abstract}
One of the major problems in forensic toolmark examination is the
same-source problem, i.e.~the identification of whether the marks on two
different objects were made by the same tool. Profilometers and confocal
microscopy and other technological advances allow us to measure 3D
surfaces in previously unforeseen resolution and also enable digitized
images. The literature provides various methods to measure similarity of
toolmarks. In this paper, we are investigating the applicability of the
Chumbley scoring method by \citet{hadler}, which was introduced for the
assessment of marks made by screwdrivers, to assess striation marks on
bullet lands for same-source identification. We provide methods to
identify parameters that minimize the error rates for matching of LEAs
at the example of the Hamby data, sets 44 and 252 measured by NIST and
CSAFE. We provide a remedial algorithm to alleviate the problem of
failed tests in the method proposed by \citet{hadler}. This increases
the power of the test and reduces error rates in the process. Type II
error rates for the Hamby data using the adjusted method showed an
improvement of more than 20\% points over the performance of the
algorithm proposed by \citet{hadler} (for a Type I error of 0.05).
Overall, type II error rates of the adjusted method are on average at
0.22, putting the method on similar footing as other single feature
matching approaches in the literature.
\end{abstract}

\noindent%
{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, land engraved areas (LEAs), algorithm
\vfill

@@

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & LEAs \\
Consecutive Matching Bullets Striae (CMS) &  &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  & &  6.26\% & 35.42\%  \\ 
Average Distance (D) &   &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & LEAs \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source land-to-land analysis reported in the literature.}
\end{table}

@@

\hypertarget{background}{%
\section{Background}\label{background}}

@@

\hypertarget{testing-setup}{%
\section{Testing setup}\label{testing-setup}}

\hypertarget{the-data}{%
\subsection{The Data}\label{the-data}}

@@

\hypertarget{setup}{%
\subsection{Setup}\label{setup}}

@@

\section{Type 2 errors: CS1 vs CS2}

\autoref{type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.

>>>>

\bigskip
\begin{abstract}
The same-source problem remains a major challenge in forensic toolmark
and firearm examination. Technological advances in surface imaging allow
measurements of 3D surfaces at previously unforeseen resolutions and
enable digitized imaging. Here, we investigate the applicability of the
Chumbley scoring method \citep{hadler}, developed for screwdriver
markings, for same-source identification of bullet striae. We provide
methods to identify parameters that minimize the error rates for
matching of LEAs using Hamby datasets 44 and 252 measured by NIST and
CSAFE. We suggest a remedial algorithm to alleviate the problem of
failed tests in the method, increasing both the power of the test and
reducing error rates. Type II error rates of the proposed method improve
by more than one third (Type I error of 0.05) and are on average about
0.22. This puts the proposed method on similar footing as other single
feature matching approaches in the literature.
\end{abstract}

\noindent%
{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, land engraved areas (LEAs), algorithm, signatures, same-source problem
\vfill

@@

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & LEAs \\
Consecutive Matching Bullets Striae (CMS) &  &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  & &  6.26\% & 35.42\%  \\ 
Average Distance (D) &   &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & LEAs \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
Cross-Correlation Function (CCF) &  &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.}
\end{table}

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

@@

\hypertarget{testing-setup}{%
\subsection{Testing setup}\label{testing-setup}}

\hypertarget{the-data}{%
\subsubsection{The Data}\label{the-data}}

@@

\hypertarget{setup}{%
\subsubsection{Setup}\label{setup}}

@@

\newpage
\section{Type 2 errors: CS1 vs CS2}

\autoref{fig:type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.

\begin{figure}[h]
